---
title: "ESD-Data-Analysis"
author: "Alex Burzinski"
date: "10/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import libraries
```{r}
library(dplyr)
```

### Import the Data
```{r}
library(RMySQL)

mySQLConn <- DBI::dbConnect(MySQL(), user='greenwic_nwu_147', password=rstudioapi::askForPassword('Database Password:')
                 ,dbname='greenwic_edu_nwu', host='67.225.186.169')

#dbListTables(mydb)

masterTable = 'greenwich_master_nwu_fall_2018_proj1'
roleTable = 'greenwich_role_nwu_fall_2018_proj1'
tagsTable = 'greenwich_tags_nwu_fall_2018_proj1'


masterData <- dbReadTable(mySQLConn, masterTable)
roleData <- dbReadTable(mySQLConn, roleTable)
tagsData <- dbReadTable(mySQLConn, tagsTable)

dbDisconnect(mySQLConn)
```

### Save data to csv files
```{r}
write.csv(masterData, file = './masterData.csv', row.names = FALSE)
write.csv(roleData, file = './roleData.csv', row.names = FALSE)
write.csv(tagsData, file = './tagsData.csv', row.names = FALSE)

write.csv(rolesCollapsed, file = './rolesCollapsed1.csv', row.names = FALSE)
```


### Read the data from files
```{r}
masterData <- read.csv('./masterData.csv', header = TRUE, stringsAsFactors = FALSE)
roleData <- read.csv('./roleData.csv', header = TRUE, stringsAsFactors = FALSE)
tagsData <- read.csv('./tagsData.csv', header = TRUE, stringsAsFactors = FALSE)
```


### Data Cleansing
```{r}
masterData$salary <- as.numeric(masterData$salary)
masterData$post_date <- as.Date(masterData$post_date, format = '%Y-%m-%d')
masterData$fill_date <- as.Date(masterData$fill_date, format = '%Y-%m-%d')
```


### Examine demand for talent for each industry
```{r}
barplot(table(masterData$vertical))

masterData$salary <- as.numeric(masterData$salary)

masterData %>%
  na.omit(cols = 'Salary') %>%
  group_by(vertical) %>%
  summarize(meanSalary = mean(salary))

masterData %>%
  na.omit(cols = time_to_fill) %>%
  group_by(vertical) %>%
  summarize(medianTimeToFill = median(time_to_fill))

masterData %>%
  group_by(vertical) %>%
  summarize(pctOpen = mean(100 * is.na(time_to_fill)))

masterData %>%
  filter(as.Date('2018-07-30') - post_date < 30) %>%
  group_by(vertical) %>%
  summarize(pctOpen = mean(100 * is.na(time_to_fill)))

masterData %>%
  group_by(vertical) %>%
  summarize(medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE))
```


### How are skills shaping demand
```{r}
masterData %>%
  filter(vertical == 'Construction') %>%
  inner_join(tagsData, by = c('job_id'), suffix = c('_master', '_tags')) %>%
  group_by(tag) %>%
  summarize(tagCount = n(), medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE)) %>%
  arrange(desc(tagCount))
```


### How metrics define high demand jobs
```{r}
masterData %>%
  filter(vertical == 'Construction') %>%
  inner_join(roleData, by = c('job_id'), suffix = c('_master', '_role')) %>%
  group_by(role) %>%
  summarize(roleCount = n(), medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE)) %>%
  arrange(desc(roleCount))
```


### How should demand be defined?
```{r}
masterData %>%
  filter(vertical == 'Construction') %>%
  summarize(medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE))
```



### How demand varies across markets?
```{r}
masterData %>%
  filter(vertical == 'Construction') %>%
  inner_join(roleData, by = c('job_id'), suffix = c('_master', '_role')) %>%
  group_by(state) %>%
  summarize(roleCount = n(), medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE)) %>%
  arrange(desc(roleCount))
```


### What are the key skills that are driving the highest demand for selected jobs?
Is this question different than 2?
```{r}
masterData %>%
  filter(vertical == 'Construction') %>%
  inner_join(tagsData, by = c('job_id'), suffix = c('_master', '_tags')) %>%
  group_by(tag) %>%
  summarize(tagCount = n(), medianTimeToFill = median(time_to_fill, na.rm = TRUE),
            pctOpen = mean(100 * is.na(time_to_fill)), medianSalary = median(salary, na.rm = TRUE)) %>%
  arrange(desc(tagCount))
```


### How do these demand drivers vary across markets?



```{r}
masterData$salary <- ifelse(is.na(masterData$salary),0,masterData$salary)
masterData$time_to_fill <- ifelse(is.na(masterData$time_to_fill),0,masterData$time_to_fill)
cor(masterData$time_to_fill, masterData$salary)
```


### Clustering
```{r}
library(tm)
library(dbscan)
library(proxy)
library(dplyr)

constructionJobs <- masterData %>%
  filter(vertical == 'Construction') %>%
  na.omit()

rolesCollapsed <- roleData %>%
  group_by(job_id) %>%
  summarize(roles = paste(role, collapse = ' '))

tagsCollapsed <- tagsData %>%
  group_by(job_id) %>%
  summarize(tags = paste(tag, collapse = ' '))

joinedData <- constructionJobs %>%
  inner_join(rolesCollapsed, by = c('job_id')) %>%
  #inner_join(tagsCollapsed, by = c('job_id')) %>%
  mutate(salary = (salary - mean(salary)) / sd(salary)) %>%
  select(time_to_fill, company, salary, latitude, longitude, location, roles)

roles_tfidf_matrix <- VectorSource(joinedData$roles) %>%
  Corpus() %>%
  DocumentTermMatrix() %>%
  weightTfIdf() %>%
  removeSparseTerms(.999) %>%
  as.matrix(weighted_tdm)

roles_kmeans <- kmeans(roles_tfidf_matrix, centers = 5)
joinedData$roleCluster <- roles_kmeans$cluster

location_dbscan <- dbscan(select(joinedData, latitude, longitude), eps = .15, minPts = 100)
location_kmeans <- kmeans(select(joinedData, latitude, longitude), centers = 10)
joinedData$locationCluster <- location_kmeans$cluster

groups  <- joinedData %>% filter(locationCluster != 0)
noise  <- joinedData %>% filter(locationCluster == 0)

ggplot(joinedData, aes(x = longitude, y = latitude, alpha = 0.5)) + 
  #geom_point(aes(fill = "grey"), noise) +
  geom_point(aes(colour = as.factor(locationCluster)), groups,
             size = 3) +
  coord_map() +
  #theme_stripped +
  #theme_empty +
  theme(legend.position = "none")

joinedData$roleCluster <- factor(joinedData$roleCluster)
joinedData$locationCluster <- factor(joinedData$locationCluster)

lm.construction <- lm(time_to_fill ~ salary + roleCluster + locationCluster, data = joinedData)
```
